---
title: "lesson2_group_assignment"
author: "Mara Alexeev"
date: "9/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BMI 713/BMIF 313qc Lecture 2: Program Flow - Group Exercises
Ted Feldman

## Your Turn: Putting it all together
Age categorization is an essential part of biomedical informatics as most predictive modeling, be it in biology, medicine or public health uses age as a covariate.

* Pick an age cohort of interest: - Infants - Pediatric and adolescents - Adults - Senior Citizens
* Work with your group to decide on an algorithm using the program flow elements we discussed today (Conditions, Creating Functions, apply() functions and subsetting)
* Find the number of patients who are classified as being part of that age group in the Neiss Package. Store the total number of remaining patients in a variable, n_other_groups
* For our purposes use the following definitions: 
  + Infants are those patients who are less than one-year-old WHO
  + Pediatric and adolescent patients are those who are ages 1 to less than 20 years of age [WHO]
  + Adult patients are ages 20 to less than 65 years of age [WHO/CMS]
  + Senior citizens are those patients who are 65 years of age or older [WHO/CMS]
  
Use two methods to demonstrate your code works correctly, including one plot

Report the time it took your code to run by placing the command ptm <- proc.time() at the start of your code to start a stopwatch command & the command: proc.time() - ptm at the end of your code to stop the clock.

Be prepared to discuss how you would demonstrate

1.    your algorithm yields optimal in terms of computational time or 
2.    how you would optimize your approach in terms of computational time.

## Your Turn: Assignment Groups
Group 1: Do this using conditionals and a for loop() and writing functions as necessary
Group 2: Do this by writing function(s) and using one of the apply() functions
Group 3: Do this by writing function(s) and using subsetting
Group 4: Do this using subsetting
Group 5: Do this by randomly sampling 10000 of the NEISS records with set.seed(5)

## Work

### Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = F)

library(ggplot2)
library(tidyverse)
library(gridExtra)
library(lubridate)
library(neiss)
library(reshape2)
```

## Start the clock!

### Load data
```{r}

data(injuries)
```

### Cohort selection with subsetting

```{r}
ptm <- proc.time()

ages <-injuries$age                               

age_infants <- ages[ages < 1] 

n_infants <- length(age_infants)
n_infants

n_other_groups <- length(ages) - n_infants
n_other_groups


proc.time() - ptm
```
### Confirmation with subsetting
```{r}
hist(age_infants, xlab="Age", ylab="Number of cases")
boxplot(age_infants)
summary(age_infants)

```

### Cohort selection with conditional

```{r}
ptm_1 <- proc.time()
ages_conditional <- injuries$age                                #extract data
less_than_age_1 <- ifelse(0<= ages_conditional & ages_conditional < 1 , yes = 1, no = 0)
n_less_than_age_1 <- sum(less_than_age_1)
n_less_than_age_1
n_other_groups_conditional <- length(ages_conditional) - n_less_than_age_1
proc.time() - ptm_1
```

### Cohort selection with function
```{r}

test_data <- injuries$age  

select_cohort <- function(data, y, z = 200 ) {
  cohort_of_interest <- as.matrix(ifelse(y <= data & data < z, yes = 1, no = 0))
  results <- apply(cohort_of_interest,2,sum)
    return(results)
}

test_function <- select_cohort(test_data, y = 0, z = 1)
test_function

peds_cohort <- select_cohort(test_data, 1, 20)
peds_cohort

```
```{r}
ptm_2 <- proc.time()
select_cohort(test_data, 0, 1)
proc.time() - ptm_2
```

### Random sampling
```{r}
ptm_3 <- proc.time()
test_data <-injuries$age
sample_size <- 10000
set.seed(5)                                         #so we all get the same case
large_sample <- test_data[sample(1:length(test_data),sample_size,replace=TRUE)]   #randomly pick a case 

large_sample_infant <- select_cohort(large_sample, 0, 1)
estimate <- large_sample_infant * length(test_data)/sample_size
estimate
proc.time() - ptm_3
```

